---
title: "Smart Meters -- Smart Behaviors? An Analysis of London Household Energy Use in Response to Dynamic Time-of-Use Tariffs"
subtitle: "ENV 797: Time Series Analysis Final Project" 
author: Maeve Gualtieri-Reed, Sylvia Hipp, Sadie Kasten, Josh Salzberg
date: April 24, 2026
output: pdf_document
editor_options: 
  chunk_output_type: console
---

\newpage
\tableofcontents 
\newpage
\listoftables 
\listoffigures 
\newpage


```{r Load Packages, message=FALSE, warning=FALSE, include=FALSE}
library(here)
library(tseries)
library(tidyverse)
library(lubridate)
library(cowplot)
library(forecast)
library(gt)
library(readxl)
library(knitr)
library(kableExtra)
```

```{r Set global theme, include=FALSE}
# set theme for formatting consistency
global_theme <- theme_classic() +
  theme(plot.title = element_text(size = 11, hjust = 0.5, face = "bold"), #Adjust title
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        axis.text.x = element_text(size = 9), # Adjust x-axis values
        axis.text.y = element_text(size = 9), # Adjust y-axis values
        axis.title.x = element_text(size = 10.5), # Adjust x-axis title
        axis.title.y = element_text(size = 10.5), # Adjust y-axis title
        legend.position = "bottom", # Define legend position
        legend.text = element_text(size = 9), # Define legend entry sizes
        legend.title = element_text(size = 9)) # Define legend name sizes

# set theme globally
theme_set(global_theme)

# **Add to any Kables**: 
# row_spec(0, align = "c", bold = TRUE) %>% 
#   kable_classic() 
```

# Introduction




# Dataset Information

## Raw Data

```{r Read Raw Data, message=FALSE, warning=FALSE, include=FALSE}
tariffs_raw <- read_excel(here("data/raw/Tariffs.xlsx"))

household_load_raw <- read.csv(here("data/raw/ExampleHouseholdLoad.csv")) %>% 
  filter(str_detect(DateTime, "^2013-01-01")) %>% 
  rename("kWhPerHalfHour" = 4)
```

The dataset for this project was obtained from a study done by UK Power Networks titled SmartMeter Energy Consumption Data in London Households.[^UKPowerNetworks] The dataset includes energy consumption readings every half hour over the period November 2011 to February 2014. The purpose of the study was to understand how price signals could support higher renewable penetration as well as reduce stress on local distribution grids. 5,566 households participated in the study and were selected as a balanced sample to represent the electricity usage of the Greater London population. 

1,123 of the participating households were subject to Dynamic Time of Use (dToU) energy prices throughout 2013. The prices were given to the households a day ahead and were displayed on the customers' smart meter or texted to them directly. Tariffs ranged from High (67.20p/kWh), Normal (11.76p/kWh), or Low (3.99p/kWh). The remaining 4,443 customers ("Standard" households) were given a flat rate tariff of 14.228p/kWh throughout the entire study period. A glimpse of the raw load data by household is shown in **[[Table XX]]** below, where `LCLid` is the unique household i.d., and `stdorTou` indicates whether the households were given standard tariff rates or subject to dToU tariffs. A glimpse of the raw data containing the tariff schedule throughout 2013 is shown in **[[Table YY]]**. 

```{r Table XX, echo=FALSE, message=FALSE, warning=FALSE, }
# household load
kable(household_load_raw %>% head(n=5), 
      caption = "Household Load Raw Data") %>% 
  row_spec(0, align = "c", bold = TRUE) %>% 
  kable_classic()

# tariffs
kable(tariffs_raw %>% head(n=5), 
      caption = "Tariff Schedule Raw Data") %>% 
  row_spec(0, align = "c", bold = TRUE) %>% 
  kable_classic()
```


[^UKPowerNetworks]: UK Power Networks. (2015). *SmartMeter energy consumption data in London households.* London Datastore. https://data.london.gov.uk/dataset/smartmeter-energy-consumption-data-in-london-households-vqm0d/.


## Data Wrangling

Both the household load and tariff schedule data were aggregated to the hour, as half-hour increments complicated the time series seasonality without adding much incremental benefits to characterizing household behavior. In cases where the tariff level differed between the top and bottom of the hour, the following logic was applied to infer tariff level for the whole hour: 

1) High and Normal -> High
2) Low and Normal -> Low
3) High and Low -> Normal

Household load was given as kWh per half hour for each household. To aggregate at the hourly level, 


[[Aggregating to the hour for both tariff dataset and household load dataset]][^PythonNote]

[[Removal of missing data + restrict to only the two whole years + separating into Std / ToU customers]]

[[K-Means Clustering]]

[^PythonNote]: The initial dataset of household load included over 200 million observations representing 2+ years of data at half-hour increments for over 5,500 households. To make the data more manageable in R, the data was initially wrangled in Python then exported to R. The code for this initial data wrangling can be found in `~/scripts/build/00_lcl_data_build.ipynb`.  


# Analysis

## Methods 



## Models



# Summary and Conclusions



_____________________________
**From Canvas**: Report Sections: 

#### Introduction, Motivation, Relevance, Objectives

Write a few paragraphs detailing the rationale for your study. This should include both the context of the topic as well as a rationale for your choice of dataset (reason for location, variables, etc.). You may choose to include citations if you like or any other reference you may have used during the project (optional).

#### Dataset information [**Sylvia**]

Provide information on how the dataset for this analysis were collected (source), the data contained in the dataset (format). Describe how you wrangled/processed your dataset to get the time series object.

Add a table that summarizes your data structure (variables, units, ranges and/or central tendencies, data source if multiple are used, etc.). This table should inserted as a `kable` function in an R chunk. Just show the first 10 rows of your data. Do not include the code used to generate your table.


#### Analysis (Methods and Models)

Describe the analysis and tests that were performed. Described the components of the time series you identified. List any packages and functions used. Include visualizations of your dataset (i.e. time series plot, ACF, PACF, etc). 

Format your R chunks so that graphs are displayed but code is not displayed. Accompany these graphs with text sections that describe the visualizations and provide context for further analyses.

Each figure should be accompanied by a caption, and referenced within the text if applicable.

#### Summary and Conclusions

Summarize your major findings from your analyses in a few paragraphs and plots. What conclusions do you draw from your findings? Any insights on how to improve the model?